---
title: "Classification Modeling"
author: "Choonghyun Ryu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Classification Modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r environment, echo = FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "")
options(tibble.print_min = 6L, tibble.print_max = 6L, width = 80)

library(alookr)
library(ranger)
library(randomForest)
```

## Preface
Once the data set is ready for model development, the model is fitted, predicted and evaluated in the following ways:

* Cleansing the dataset
* Split the data into a train set and a test set
* **Modeling and Evaluate, Predict**
    + **Modeling**
        + **Binary classification modeling**
    + **Evaluate the model**
        + **Predict test set using fitted model**
        + **Calculate the performance metric**
        + **Plot the ROC curve** 
        + **Tunning the cut-off**         
    + **Predict**
        + **Predict** 
        + **Predict with cut-off**  
        
The alookr package makes these steps fast and easy:

## Data: Wisconsin Breast Cancer Data
`BreastCancer` of `mlbench package` is a breast cancer data. The objective is to identify each of a number of benign or malignant classes. 

A data frame with 699 observations on 11 variables, one being a character variable, 9 being ordered or nominal, and 1 target class.:

* `Id` : character.	Sample code number
* `Cl.thickness` : ordered factor. Clump Thickness
* `Cell.size` :	ordered factor. Uniformity of Cell Size
* `Cell.shape` : ordered factor. Uniformity of Cell Shape
* `Marg.adhesion` :	ordered factor. Marginal Adhesion
* `Epith.c.size` : ordered factor. Single Epithelial Cell Size
* `Bare.nuclei` :	factor. Bare Nuclei
* `Bl.cromatin` :	factor. Bland Chromatin
* `Normal.nucleoli` :	factor. Normal Nucleoli
* `Mitoses` :	factor. Mitoses
* `Class` :	factor. Class. level is `benign` and `malignant`.

```{r load_data}
library(mlbench)
data(BreastCancer)

# class of each variables
sapply(BreastCancer, function(x) class(x)[1])
```

## Preperation the data
Perform data preprocessing as follows.:

* Find and imputate variables that contain missing values.
* Split the data into a train set and a test set.
* To solve the imbalanced class, perform sampling in the train set of raw data.
* Cleansing the dataset for classification modeling.

### Fix the missing value with `dlookr::imputate_na()`
find the variables that include missing value. and imputate the missing value using imputate_na() in dlookr package.

```{r imputate_data, message=FALSE, warning=FALSE}
library(dlookr)
library(dplyr)

# variable that have a missing value
diagnose(BreastCancer) %>%
  filter(missing_count > 0)

# imputation of missing value
breastCancer <- BreastCancer %>%
  mutate(Bare.nuclei = imputate_na(BreastCancer, Bare.nuclei, Class,
                         method = "mice", no_attrs = TRUE, print_flag = FALSE))
```

## Split data set
### Splits the dataset into a train set and a test set with `split_by()`

`split_by()` in the alookr package splits the dataset into a train set and a test set.

The ratio argument of the `split_by()` function specifies the ratio of the train set.

`split_by()` creates a class object named split_df.

```{r split_data, warning=FALSE}
library(alookr)

# split the data into a train set and a test set by default arguments
sb <- breastCancer %>%
  split_by(target = Class)

# show the class name
class(sb)

# split the data into a train set and a test set by ratio = 0.6
tmp <- breastCancer %>%
  split_by(Class, ratio = 0.6)
```

The `summary()` function displays the following useful information about the split_df object:

* random seed : The random seed is the random seed used internally to separate the data
* split data : Information of splited data
    + train set count : number of train set 
    + test set count : number of test set
* target variable : Target variable name
    + minority class : name and ratio(In parentheses) of minority class
    + majority class : name and ratio(In parentheses) of majority class

```{r split_summary, warning=FALSE}
# summary() display the some information
summary(sb)

# summary() display the some information
summary(tmp)
```

### Check missing levels in the train set

In the case of categorical variables, when a train set and a test set are separated, a specific level may be missing from the train set.

In this case, there is no problem when fitting the model, but an error occurs when predicting with the model you created. Therefore, preprocessing is performed to avoid missing data preprocessing.

In the following example, fortunately, there is no categorical variable that contains the missing levels in the train set.

```{r split_check, warning=FALSE}
# list of categorical variables in the train set that contain missing levels
nolevel_in_train <- sb %>%
  compare_target_category() %>% 
  filter(is.na(train)) %>% 
  select(variable) %>% 
  unique() %>% 
  pull

nolevel_in_train

# if any of the categorical variables in the train set contain a missing level, 
# split them again.
while (length(nolevel_in_train) > 0) {
  sb <- breastCancer %>%
    split_by(Class)

  nolevel_in_train <- sb %>%
    compare_target_category() %>% 
    filter(is.na(train)) %>% 
    select(variable) %>% 
    unique() %>% 
    pull
}
```

## Handling the imbalanced classes data with `sampling_target()`
### Issue of imbalanced classes data
Imbalanced classes(levels) data means that the number of one level of the frequency of the target variable is relatively small.
In general, the proportion of positive classes is relatively small. For example, in the model of predicting spam, the class of interest spam is less than non-spam.

Imbalanced classes data is a common problem in machine learning classification.

`table()` and `prop.table()` are traditionally useful functions for diagnosing imbalanced classes data. However, alookr's `summary()` is simpler and provides more information.

```{r show_ratio, warning=FALSE}
# train set frequency table - imbalanced classes data
table(sb$Class)

# train set relative frequency table - imbalanced classes data
prop.table(table(sb$Class))

# using summary function - imbalanced classes data
summary(sb)
```

### Handling the imbalanced classes data

Most machine learning algorithms work best when the number of samples in each class are about equal. And most algorithms are designed to maximize accuracy and reduce error. So, we requre handling an imbalanced class problem.

sampling_target() performs sampling to solve an imbalanced classes data problem.

### Resampling - oversample minority class
Oversampling can be defined as adding more copies of the minority class.

Oversampling is performed by specifying "ubOver" in the method argument of the `sampling_target()` function.

```{r sampling_over, warning=FALSE}
# to balanced by over sampling
train_over <- sb %>%
  sampling_target(method = "ubOver")

# frequency table 
table(train_over$Class)
```

### Resampling - undersample majority class
Undersampling can be defined as removing some observations of the majority class.

Undersampling is performed by specifying "ubUnder" in the method argument of the `sampling_target()` function.

```{r sampling_under, warning=FALSE}
# to balanced by under sampling
train_under <- sb %>%
  sampling_target(method = "ubUnder")

# frequency table 
table(train_under$Class)
```

### Generate synthetic samples - SMOTE
SMOTE(Synthetic Minority Oversampling Technique) uses a nearest neighbors algorithm to generate new and synthetic data. 

SMOTE is performed by specifying "ubSMOTE" in the method argument of the `sampling_target()` function.

```{r sampling_smote, warning=FALSE}
# to balanced by SMOTE
train_smote <- sb %>%
  sampling_target(seed = 1234L, method = "ubSMOTE")

# frequency table 
table(train_smote$Class)
```

## Cleansing the dataset for classification modeling with `cleanse()`
The `cleanse()` cleanse the dataset for classification modeling.

This function is useful when fit the classification model. This function does the following.:

* Remove the variable with only one value. 
* And remove variables that have a unique number of values relative to the number of observations for a character or categorical variable. 
    + In this case, it is a variable that corresponds to an identifier or an identifier. 
* And converts the character to factor.

In this example, The `cleanse()` function removed a variable ID with a high unique rate.

```{r clean_data, warning=FALSE}
# clean the training set
train <- train_smote %>%
  cleanse
```

## Extract test set for evaluation of the model with `extract_set()`

```{r extract_test, warning=FALSE}
# extract test set
test <- sb %>%
  extract_set(set = "test")
```

## Binary classification modeling with `run_models()`
`run_models()` performs some representative binary classification modeling using `split_df` object created by `split_by()`.

`run_models()` executes the process in parallel when fitting the model. However, it is not supported in MS-Windows operating system and RStudio environment.

Currently supported algorithms are as follows.:

* logistic : logistic regression using `stats` package
* rpart : Recursive Partitioning Trees using `rpart` package
* ctree : Conditional Inference Trees using `party` package
* randomForest :Classification with Random Forest using `randomForest` package
* ranger : A Fast Implementation of Random Forests using `ranger` package
* xgboost : Extreme Gradient Boosting using `xgboost` package

`run_models()` returns a `model_df` class object.

The `model_df` class object contains the following variables.:

* step : character. The current stage in the classification modeling process.
    + For objects created with `run_models()`, the value of the variable is "1.Fitted".
* model_id : model identifiers
* target : name of target variable
* positive : positive class in target variable
* fitted_model : list. Fitted model object by model_id's algorithms


```{r fit_model, message=FALSE, warning=FALSE}
result <- train %>% 
  run_models(target = "Class", positive = "malignant")
result
```

## Evaluate the model
Evaluate the predictive performance of fitted models.

### Predict test set using fitted model with `run_predict()`

`run_predict()` predict the test set using `model_df` class fitted by `run_models()`.

`run_predict ()` is executed in parallel when predicting by model. However, it is not supported in MS-Windows operating system and RStudio environment.

The `model_df` class object contains the following variables.:

* step : character. The current stage in the classification modeling process.
    + For objects created with `run_predict()`, the value of the variable is "2.Predicted".
* model_id : character. Type of fit model.
* target : character. Name of target variable.
* positive : character. Level of positive class of binary classification.
* fitted_model : list. Fitted model object by model_id's algorithms.
* predicted : result of predcit by each models

```{r predict}
pred <- result %>%
  run_predict(test)
pred
```

### Calculate the performance metric with `run_performance()`

`run_performance()` calculate the performance metric of `model_df` class predicted by `run_predict()`.

`run_performance ()` is performed in parallel when calculating the performance evaluation metrics However, it is not supported in MS-Windows operating system and RStudio environment.

The `model_df` class object contains the following variables.:

* step : character. The current stage in the classification modeling process.
    + For objects created with `run_performance()`, the value of the variable is "3.Performanced".
* model_id : character. Type of fit model.
* target : character. Name of target variable.
* positive : character. Level of positive class of binary classification.
* fitted_model : list. Fitted model object by model_id's algorithms
* predicted : list. Predicted value by individual model. Each value has a predict_class class object.
* performance : list. Calculate metrics by individual model. Each value has a numeric vector.

```{r performance1}
# Calculate performace metrics.
perf <- run_performance(pred)
perf
```

The performance variable contains a list object, which contains 15 performance metrics:

* ZeroOneLoss : Normalized Zero-One Loss(Classification Error Loss).
* Accuracy : Accuracy.
* Precision : Precision.
* Recall : Recall.
* Sensitivity : Sensitivity.
* Specificity : Specificity.
* F1_Score : F1 Score.
* Fbeta_Score : F-Beta Score.
* LogLoss : Log loss / Cross-Entropy Loss.
* AUC : Area Under the Receiver Operating Characteristic Curve (ROC AUC).
* Gini : Gini Coefficient.
* PRAUC : Area Under the Precision-Recall Curve (PR AUC).
* LiftAUC : Area Under the Lift Chart.
* GainAUC : Area Under the Gain Chart.
* KS_Stat : Kolmogorov-Smirnov Statistic.

```{r performance2}
# Performance by analytics models
performance <- perf$performance
names(performance) <- perf$model_id
performance
```

If you change the list object to tidy format, you'll see the following at a glance:

```{r performance3}
# Convert to matrix for compare performace.
sapply(performance, "c")
```

`compare_performance()` return a list object(results of compared model performance). and list has the following components:

* recommend_model : character. The name of the model that is recommended as the best among the various models.
* top_count : numeric. The number of best performing performance metrics by model.
* mean_rank : numeric. Average of ranking individual performance metrics by model.
* top_metric : list. The name of the performance metric with the best performance on individual performance metrics by model.

In this example, `compare_performance()` recommend the **"ranger"** model.

```{r compare_performance}
# Compaire the Performance metrics of each model
comp_perf <- compare_performance(pred)
comp_perf
```

### Plot the ROC curve with `plot_performance()`

`compare_performance()` plot ROC curve.

```{r ROC, fig.height=4, fig.width=7}
# Plot ROC curve
plot_performance(pred)
```


### Tunning the cut-off

In general, if the prediction probability is greater than 0.5 in the binary classification model, it is predicted as `positive class`.
In other words, 0.5 is used for the cut-off value.
This applies to most model algorithms. However, in some cases, the performance can be tuned by changing the cut-off value.

`plot_cutoff ()` visualizes a plot to select the cut-off value, and returns the cut-off value.

```{r cutoff, warning=FALSE, fig.height=4, fig.width=7}
pred_best <- pred %>% 
  filter(model_id == comp_perf$recommend_model) %>% 
  select(predicted) %>% 
  pull %>% 
  .[[1]] %>% 
  attr("pred_prob")

cutoff <- plot_cutoff(pred_best, test$Class, "malignant", type = "mcc")
cutoff

cutoff2 <- plot_cutoff(pred_best, test$Class, "malignant", type = "density")
cutoff2

cutoff3 <- plot_cutoff(pred_best, test$Class, "malignant", type = "prob")
cutoff3
```

### Performance comparison between prediction and tuned cut-off with `performance_metric()`

Compare the performance of the original prediction with that of the tuned cut-off. 
Compare the cut-off with the non-cut model for the model with the best performance `comp_perf$recommend_model`.

```{r predit_cutoff}
comp_perf$recommend_model

# extract predicted probability
idx <- which(pred$model_id == comp_perf$recommend_model)
pred_prob <- attr(pred$predicted[[idx]], "pred_prob")

# or, extract predicted probability using dplyr
pred_prob <- pred %>% 
  filter(model_id == comp_perf$recommend_model) %>% 
  select(predicted) %>% 
  pull %>% 
  "[["(1) %>% 
  attr("pred_prob")

# predicted probability
pred_prob  

# compaire Accuracy
performance_metric(pred_prob, test$Class, "malignant", "Accuracy")
performance_metric(pred_prob, test$Class, "malignant", "Accuracy",
                   cutoff = cutoff)

# compaire Confusion Matrix
performance_metric(pred_prob, test$Class, "malignant", "ConfusionMatrix")
performance_metric(pred_prob, test$Class, "malignant", "ConfusionMatrix", 
                   cutoff = cutoff)

# compaire F1 Score
performance_metric(pred_prob, test$Class, "malignant", "F1_Score")
performance_metric(pred_prob, test$Class,  "malignant", "F1_Score", 
                   cutoff = cutoff)
performance_metric(pred_prob, test$Class,  "malignant", "F1_Score", 
                   cutoff = cutoff2)
```

If the performance of the tuned cut-off is good, use it as a cut-off to predict positives.

## Predict
If you have selected a good model from several models, then perform the prediction with that model.

### Create data set for predict
Create sample data for predicting by extracting 100 samples from the data set used in the previous under sampling example.

```{r predit_data, warning=FALSE}
data_pred <- train_under %>% 
  cleanse 

set.seed(1234L)
data_pred <- data_pred %>% 
  nrow %>% 
  seq %>% 
  sample(size = 50) %>% 
  data_pred[., ]
```

### Predict with alookr and dplyr
Do a predict using the `dplyr` package. The last `factor()` function eliminates unnecessary information.

```{r predit_final, warning=FALSE}
pred_actual <- pred %>%
  filter(model_id == comp_perf$recommend_model) %>% 
  run_predict(data_pred) %>% 
  select(predicted) %>% 
  pull %>% 
  "[["(1) %>% 
  factor()

pred_actual
```

If you want to predict by cut-off, specify the `cutoff` argument in the `run_predict()` function as follows.:

In the example, there is no difference between the results of using cut-off and not.

```{r predit_final2, warning=FALSE}
pred_actual2 <- pred %>%
  filter(model_id == comp_perf$recommend_model) %>% 
  run_predict(data_pred, cutoff) %>% 
  select(predicted) %>% 
  pull %>% 
  "[["(1) %>% 
  factor()

pred_actual2

sum(pred_actual != pred_actual2)
```

